---
title: "R_movie_ratings"
author: "Marlon Nurse"
date: "2026-09-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Connect Data / CSV files 

```{r}
ratings_df <- read.csv(
  "https://raw.githubusercontent.com/mj-nurse/mnurse-data_607/refs/heads/main/assignment-02a-movie-ratings/movie_ratings.csv",
  na.strings = c("", "NULL")
)

head(ratings_df)
```


## Calculating Results 
```{r}
library(dplyr)

movie_summary <- ratings_df %>%
  group_by(title) %>%
  summarise(
    rating_count = sum(!is.na(rating)),
    missing_ratings = sum(is.na(rating)),
    average_rating = round(mean(rating, na.rm = TRUE), 2)
  ) %>%
  arrange(desc(average_rating))

movie_summary
```

## Conclusion 

Participants who selected "Not Seen" were stored as missing values rather than assigning them a numeric rating. In R, these values are represented as `NA`. When calculating average ratings, I used only actual ratings from 1 through 5 were included in the calculation. This prevents a movie from receiving an artificially low score simply because a participant had not seen it. So in conclusion of my data, the people voted The Odyssey as the best movie of the 6 with an average of 4.15 / 5. While, Backrooms scored the lowest on average with a score of 3.10 / 5. Fun fact, Obsession & Spider-Man were tied for the most watched/rated movie with 18 / 22 ratings. 


